---
title: "MLB EDA"
author: "Austin Stephen, Advaith Vijay, and Claire Dudley"
date: "6/14/2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 9, fig.height=3.5, fig.retina = 3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE }
library(xaringanthemer)
style_mono_accent(base_color = "#808080",
  white_color = "#FFFFFF",
  background_color = "#FFFFFF")
```

class: left, top
## MLB Batted Balls 
2021 season  
Different game conditions influence outcomes

???
All of our exploration can be put under the umbrella of how do different conditions 
in the game impact the outcome. We explore this through 3 hypothesis and a clustering 
example.

---

## This Slide Should Explain Our Data

* Every batted ball from the 2021 season 
* 94 => 35 columns about the hits 



???
Note for slide 2
---

## First Hypothesis 

---

## Second Hypothesis

---

## Third Hypothesis 

---

## Clustering example

---

## R shiny link

---
## Conclusion Slide 

---




