---
title: "MLB 2021 Batted Balls EDA"
author: "Austin, Advaith and Claire"
date: "6/14/2021"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 9, fig.height=3.5, fig.retina = 3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE }
library(xaringanthemer)
style_mono_accent(base_color = "#808080",
  white_color = "#FFFFFF",
  background_color = "#fff7e6",
  base_font_size = "24px")
```

class: left, top
## MLB Batted Balls Data
* Exclusively 2021 season data 
* Exploring how different game conditions influence outcomes
* (Probably need to add more about motivation)

???
All of our exploration can be put under the umbrella of how different conditions 
in the game impact play outcome. We explore this through 3 hypothesis and a clustering 
example.


---

## The Data 

* Every batted ball from the 2021 season  

* Summaries by batter 

* Summaries by pitcher 

???
I want it to reveal 1 at a time summary table for pitcher and batters
---

## Hit Location 
* How does hit location influence bases taken?  
(insert code)
* How does batters handedness influence hit location?  
(insert code)
---

## Bases Taken
* How is hit distance distributed?  
(insert code)

* Bases taken as a function of hit distance?  
(insert code)

* Who is the most efficient base runner this year?  
(insert code)

---

## Stadiums and hitting
* Advaith fill in here!
---

## Clustering example
* Claire fill in here 

---
## Conclusion Slide 




